---
title: "Wacaha"
subtitle: "Pricing and Retail Analytics"
author: "Natalie Kvochak, Maryclare Leonard, Sean O'Grady"
date: "3/29/21"
output:
  pdf_document:
    number_sections: yes
colorlinks: true
urlcolor: 'red'
fontsize: 12pt
header-includes: \usepackage{color,soul}
---

```{r, include=FALSE}
# We recommend including this entire code chunk at the top of all 
# your R Markdown files. 'include=FALSE' omits this code in the output
# but R still runs it.

# Load libraries needed for analysis
library(knitr)                 # Creating PDF output using Markdown
library(tidyverse)             # Data manipulation and graphing
library(readxl)                # To use read_excel()
library(psych)                 # To use corr.test() for correlation matrices
theme_set(theme_bw())          # Get rid of gray background in plots

# Set some defaults for the R Markdown document
opts_chunk$set(echo=TRUE,      # Print the code in all the chunks
               warning=FALSE,  # Don't print warning statements
               message=FALSE,  # Don't print other R output messages
               comment=NA)     # Helps produce prettier output
```

# Technical section

## Data setup
```{r}
# Read in the data
df <- read_excel("small_salsa.xlsx")

# Create logged prices and quantities
df <- df %>% 
  mutate(ln_p = log(price),
         ln_q = log(units),
         Dholiday = factor(holiday),
         Dstore = factor(store),
         Dzone= factor(zone))

```

## Data Exploration
```{r}
#plot price versus units
plot(df$price, df$units,main="Units as a function of price",
     xlab="Price",ylab="Units",pch=19)

plot(df$ln_q,df$ln_p,main="Log price as a function of log units",
     xlab="Log Units", ylab="Log Price",pch=19)
```

## Correlation
```{r}
corr.test(df %>% select(price, pop, units, holiday))
```

## Regression
```{r}
reg1 <- lm(ln_q ~ ln_p + Dzone + Dholiday, data=df)
summary(reg1)
```

The regression indicates that the price elasticity is `r round(reg1$coefficients['ln_p'],3)`.

## Regression Calculation for Zone 1
```{r}
reg2 <- lm(ln_q ~ ln_p + Dholiday, data=df %>% filter(Dzone==1))
summary(reg2)
```

The regression indicates that the price elasticity  for Zone 1 is `r round(reg2$coefficients['ln_p'],3)`.

## Regression Calculation for Zone 2
``` {r}
reg3 <- lm(ln_q ~ ln_p + Dholiday, data=df %>% filter(Dzone==2))
summary(reg3)
```

The regression indicates that the price elasticity for Zone 2 is `r round(reg3$coefficients['ln_p'],3)`.

## Population Regression Calculation for Zone 1
```{r}
reg4 <- lm(ln_q ~ ln_p + pop, data=df %>% filter(Dzone==1))
summary(reg4)
```

The regression indicates that the price elasticity  for Zone 1 is `r round(reg4$coefficients['ln_p'],3)`.

# Managerial Discussion

In Scenario 1, the promotion price of £1.19, Tesco’s trade budget is £1540 in zone 1 and £599 in zone 2 during an average week. This equates to a total weekly trade budget of £2139. If the promotion price is changed to £1.09 in scenario 2, the trade budget increases to £1880 in zone 1 and £736 in zone 2. The new total weekly trade budget increases to £2616. While scenario 2 require Wahaca to pay more money to Tesco in the scanbacks, it ends up being the more profitable option for Wahaca. The high elasticity of the salsa leads to large increases in quantity demanded with decreased prices. This increase in demand is sufficient to offset the decrease in price, as shown by the increase in total gross profit from £6009 to £6105. The issue with scenario 2, though, is Tesco is making less money than they did with the original promotion price. In the original promotion, Wahaca covered £0.21 of the £0.40 decrease with their scanback. In covering the majority of the scanback, Wahaca makes less money in promoted weeks under scenario 1 than they do during regular weeks. In the switch to a price of £1.09, Wahaca still pays £0.21 per item, but the price decrease is now £0.50. This means that Tesco is stuck absorbing a larger percentage of the promotion, making the price change less profitable for them. If the scanback price were changed to £0.24 in scenario 2, Wahaca would still be making more money during a promotion week than a regular week, making it preferable to scenario 1 from their perspective. Additionally, the increase in the scanback would make the price change more profitable for Tesco as well. Given the fact that this salsa is new to the market, scenario 2 may also make more sense to push volume and increase market penetration.
