---
title: "IA4: Tropicana"
subtitle: "Section 2"
author: "Maryclare Leonard, Sean O'Grady, & Natalie Kvochak"
date: "4/5/22"
output:
  pdf_document:
    number_sections: yes
colorlinks: true
urlcolor: 'red'
fontsize: 12pt
header-includes: \usepackage{color,soul}
---

```{r, include=FALSE}
# We recommend including this entire code chunk at the top of all 
# your R Markdown files. 'include=FALSE' omits this code in the output
# but R still runs it.

# Load libraries needed for analysis
library(knitr)                 # Creating PDF output using Markdown
library(tidyverse)             # Data manipulation and graphing
library(readxl)                # To use read_excel()
library(psych)                 # To use corr.test() for correlation matrices
theme_set(theme_bw())          # Get rid of gray background in plots

# Set some defaults for the R Markdown document
opts_chunk$set(echo=TRUE,      # Print the code in all the chunks
               warning=FALSE,  # Don't print warning statements
               message=FALSE,  # Don't print other R output messages
               comment=NA)     # Helps produce prettier output
```

# Technical section

This is where your technical material should go. You might start by reading in the data.

```{r}
# Read the data
df <- read_excel("tropicana.xlsx")

# Create some variables
df <- df %>% mutate(ln_p=log(price_tr_12),
                   ln_q=log(units_tr_12),
                   Dmerch=factor(merch_tr_12),
                   Dzone=factor(zone),
                   Dyear=factor(year))
```

## Explore Your Data
You are encouraged to use these visualizations of your data, but you can do other visualizations if you desire.

```{r}

# Plot Units Sold
ggplot(df, aes(x=weeknum, y=units_tr_12)) + 
  geom_line() + 
  facet_wrap( ~ zone, ncol=2) +
  labs(x="Weeknum", y="Units")

# Plot Prices
ggplot(df, aes(x=weeknum)) +
  geom_line(aes(y=reg_tr_12), color="blue") +
  geom_line(aes(y=price_tr_12), color="red") + 
  facet_wrap( ~ zone, ncol=2) +
  labs(x="Weeknum", y="Price")

# Plot Log Price vs. Log Units
ggplot(df, aes(x=ln_q, y=ln_p)) + 
  geom_point() + 
  facet_wrap( ~ zone, ncol=2) +
  geom_smooth(method='lm', se=FALSE, orientation="y") +
  labs(x="ln(Units)", y="ln(Price)")

# Plot Log Price vs. Log Units
ggplot(df, aes(y=ln_q, x=ln_p)) + 
  geom_point() + 
  facet_wrap( ~ zone, ncol=2) +
  geom_smooth(method='lm', se=FALSE) +
  labs(y="ln(Units)", x="ln(Price)") + coord_flip()

ggplot(df, aes(y=ln_q, x=ln_p, col=Dmerch)) + 
  geom_point() + 
  facet_wrap( ~ zone, ncol=2) +
  geom_smooth(method='lm', se=FALSE) +
  labs(y="ln(Units)", x="ln(Price)") + 
  coord_flip()

df.med <- df %>% filter(zone == "Medium")

ggplot(df.med, aes(x=ln_q, y=ln_p, col=Dmerch)) + 
  geom_point() + 
  geom_smooth(method='lm', se=FALSE) +
  labs(x="ln(Units)", y="ln(Price)")



summary(lm(ln_q ~ ln_p + ln_p:Dmerch + Dmerch , data=df.med))

summary(lm(ln_q ~ ln_p, data=df.med %>% filter(merch_tr_12 == 0)))
ggplot(df.med %>% filter(merch_tr_12 == 0), aes(x=ln_q, y=ln_p)) + 
  geom_point() + 
  xlim(5,12) + ylim(-0.1, 0.8) +
  geom_smooth(method='lm', se=FALSE) +
  labs(x="ln(Units)", y="ln(Price)")

summary(lm(ln_q ~ ln_p, data=df.med %>% filter(merch_tr_12 == 1)))
ggplot(df.med %>% filter(merch_tr_12 == 1), aes(x=ln_q, y=ln_p)) + 
  geom_point() + 
  xlim(5,12) + ylim(-0.1, 0.8) +
  geom_smooth(method='lm', se=FALSE) +
  labs(x="ln(Units)", y="ln(Price)")




####
summary(lm(ln_q ~ ln_p, data=df.med %>% filter(merch_tr_12 == 0)))
p1 <- ggplot(df.med %>% filter(merch_tr_12 == 0), aes(y=ln_q, x=ln_p)) + 
  geom_point() + 
  ylim(5,12) + xlim(-0.1, 0.8) +
  geom_smooth(method='lm', se=FALSE) +
  labs(x="ln(P)", y="ln(Q)")

summary(lm(ln_q ~ ln_p, data=df.med %>% filter(merch_tr_12 == 1)))
p2 <- ggplot(df.med %>% filter(merch_tr_12 == 1), aes(y=ln_q, x=ln_p)) + 
  geom_point() + 
  ylim(5,12) + xlim(-0.1, 0.8) +
  geom_smooth(method='lm', se=FALSE) +
  labs(x="ln(P)", y="ln(Q)")

gridExtra::grid.arrange(p1,p2, ncol=2)

ggplot(df.med, aes(y=ln_q, x=ln_p, col=Dmerch)) + 
  geom_point() + 
  #xlim(5,12) + ylim(-0.1, 0.8) +
  geom_smooth(method='lm', se=FALSE) +
  coord_flip()


p1 <- ggplot(df, aes(x=ln_q, y=ln_p, col=Dmerch)) + 
  geom_point() + 
  facet_wrap( ~ zone, ncol=2) +
  geom_smooth(method='lm', se=FALSE, orientation="y") +
  labs(x="ln(Units)", y="ln(Price)") + labs(title="Original")

p2 <- ggplot(df, aes(y=ln_q, x=ln_p, col=Dmerch)) + 
  geom_point() + 
  facet_wrap( ~ zone, ncol=2) +
  geom_smooth(method='lm', se=FALSE) +
  labs(y="ln(Units)", x="ln(Price)") + 
  coord_flip() + labs(title="New")

gridExtra::grid.arrange(p1,p2, ncol=2)





```


## Run the main regression model

You are encouraged to use this regression, however you can complement your analysis with data exploration and additional regressions.

```{r}
reg.tr <- lm(ln_q ~ ln_p + ln_p:Dmerch + Dmerch + 
               Dyear + Dzone, data=df)
```

Be sure to finalize your regression and ... No more regressions to run after this!

## Scenario Analysis

### Scenario 0 (Baseline)

```{r}
#run main regression 
reg.tr <- lm(ln_q ~ ln_p + ln_p:Dmerch + Dmerch + 
               Dyear + Dzone, data=df)
#copy data 
df.pred = df

#Predict ln(units) using the actual data -- the output are the fitted values
df.pred$ln_pred_units_tr_12<- predict(reg.tr, df.pred)

# Convert Log to Units and Compute Profit
df.pred <- df.pred %>% mutate(units_tr_12 = exp(ln_pred_units_tr_12),
profit_tr_12 = units_tr_12*(price_tr_12 - cost_tr_12))

# Create a table to summarize predicted profit for baseline
df.pred %>%
  group_by(year) %>%
  summarize(sum_profits = sum(profit_tr_12)) %>%
  kable(digits=0)

#create a table to summarize by zone and filter by 2018 
df.pred %>%
  filter(year==2018)%>%
  group_by(zone) %>%
  summarize(sum_profits = sum(profit_tr_12)) %>%
  kable(digits=0)
```


### Scenario 1

```{r}
# Implement the scenario!
# This is *not* the same scenario from the Energy Bar case in class.
# You need to modify the original data to alter the price and merchandise variables
# using the *_scenario variables that we provide.

df.pred2 <- df

df.pred2 <- df.pred2 %>%
  filter(year==2018) %>%
  mutate (price_tr_12 = price_tr_12_scenario1,
           merch_tr_12 = merch_tr_12_scenario1)%>%
  mutate (ln_p = log(price_tr_12),
          Dmerch = factor(merch_tr_12))

# Use the regression model to predict using the modified data
df.pred2$ln_pred_units_tr_12 <- predict(reg.tr, df.pred2)

# Convert Log to Units and Compute Profit
df.pred2 <- df.pred2 %>% mutate(units_tr_12 = exp(ln_pred_units_tr_12),
profit_tr_12_scenario1 = units_tr_12*(price_tr_12 - cost_tr_12))

# Create a table to summarize predicted profit for the alternative scenario
df.pred2 %>%
  group_by(year) %>%
  summarize(sum_profits = sum(profit_tr_12_scenario1)) %>%
  kable(digits=0)

#create a table to summarize by zone 
df.pred2 %>%
  group_by(zone) %>%
  summarize(sum_profits = sum(profit_tr_12_scenario1)) %>%
  kable(digits=0)

```



### Compare the Scenarios

# Managerial Discussion

Managerial discussion goes here.

