
---
title: "R Review Session"
author: "Brett Gordon and Joonhyuk Yang"
date: "10/07/21"
output:
  pdf_document:
    number_sections: yes
colorlinks: true
urlcolor: 'red'
fontsize: 12pt
---

```{r, include=FALSE}
# second chunk - load required packages - don't touch, but run this part 
# click run - R run this chunk - run this first
# We recommend including this entire code chunk at the top of all 
# your R Markdown files. 'include=FALSE' omits this code in the output
# but R still runs it.

# Load libraries needed for analysis
library(knitr)                 # Creating PDF output using Markdown
library(tidyverse)             # Data manipulation and graphing
library(readxl)                # To use read_excel()
library(psych)                 # To use corr.test() for correlation matrices
theme_set(theme_bw())          # Get rid of gray background in plots

# Set some defaults for the R Markdown document
opts_chunk$set(echo=TRUE,      # Print the code in all the chunks
               warning=FALSE,  # Don't print warning statements
               message=FALSE,  # Don't print other R output messages
               comment=NA)     # Helps produce prettier output
```

# Setup

This example file uses data from Dominick's, a now defunct grocery store chain in Chicago. First, I read in the data and print out a quick summary of the variables.


```{r}
#  run line 42 - df table created
df <- read_excel("dominicks_oj.xlsx")
str(df) # compactly displaying the internal structure of a R object
head(df) # print first few lines of the data
```

# Manipulating the data

Let's create some new variables.

```{r}
df <- df %>% 
  mutate(lnp = log(price),
         lnq = log(units),
         Dmerch = factor(merch),
         Dholiday = factor(holiday),
         Dsize = factor(size),
         Dzone = factor(zone))
head(df) #append these new variables back to df
```

Quick tabulations by zone, brand, and brand-size.

```{r}
table(df$zone) #dollar sign to indicate data frame
table(df$brand)
table(df$brand, df$size) 
```

Summary statistics by brand.

```{r}
# Total sales, mean price, number of observations *for each* brand
df %>%
  group_by(brand) %>%
  summarize(sum_sales = sum(units),
            mean_price = mean(price),
            num_obs = n())

```
 
Summary statistics for a subset of the data.
 
```{r}
df %>%
  filter(size==12) %>%
  group_by(brand) %>%
  summarize(sum_sales = sum(units),
            mean_price = mean(price),
            num_obs = n())
```

Let me create some handy subsets of the data (and note my variable naming scheme).

```{r}
df.low <- df %>% filter(zone=="Low")
df.mmaid12 <- df %>% filter(brand=="MMAID" & size==12)
df.low.mmaid12 <- df.mmaid12 %>% filter(zone=="Low")
```

# Plots

## Scatter plots

```{r}
ggplot(df.low.mmaid12, aes(x=lnq, y=lnp)) +
  geom_point()

ggplot(df.low.mmaid12, aes(x=lnq, y=lnp, color=Dmerch)) +
  geom_point()

ggplot(df.low.mmaid12, aes(x=lnq, y=lnp)) +
  geom_point() + 
  facet_wrap(~ Dmerch) 

ggplot(df.low.mmaid12, aes(x=lnq, y=lnp)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE)

```

## Line plots

```{r}
ggplot(df.low.mmaid12, aes(x=week, y=price)) + 
  geom_line() + 
  labs(x="Weeknum", y="Price", title="Minute Maid 16 oz. in Zone = Medium")
```

# Regression

```{r}
reg1 <- lm(lnq ~ lnp, data=df.low.mmaid12)
summary(reg1)
```

The regression indicates that the price elasticity is `r round(reg1$coefficients['lnp'],3)`.  Now I will add zone dummies and an interaction between `lnp` and `Dmerch`.

```{r}
reg2 <- lm(lnq ~ lnp + Dzone + lnp:Dmerch, data=df.mmaid12)
summary(reg2)
```

Now the price elasticity is `r round(reg2$coefficients['lnp'],3)`, which is pretty similar to the first regression. Both estimates are within the reasonable range for a price elasticity.
